The complex network of trophic interactions in a subAntarctic oceanic Marine Protected

### 2023-07-12 MS_ToSubmit/MS_Burdwood_foodweb.Rmd
Schejter suggs.

abstract: |
  Globally, the total area of the world ocean designated under marine protection has increased in the recent decades. The majority of these Marine Protected Areas (MPAs) focus on the presence of particularly vulnerable, keystone, or charismatic species, large numbers of endemic species, and/or high biodiversity across taxonomic levels. In the sub-Antarctic region, the level of ocean protection is mainly associated to oceanic islands, except for the MPAs Namuncurá - Burdwood Bank I and II (MPA N-BB, ~53º–55ºS and ~56º–62ºW), which are associated to a submarine plateau and its southern adjacent deep slope region, respectively. Here, we present the first detailed analysis of the network of predator-prey interactions (food web) for the MPA N-BB ecosystem. We applied a network approach to characterise the food web in terms of complexity and structure, and identifying the species' role in such a framework. In terms of complexity, the MPA N-BB food web consisted of 1788 predator-prey interactions and 380 species, with a link density of 4.70 and a connectance of 0.01. In terms of structure, almost half of the consumers were omnivores (0.48), and the network displayed a small-world pattern. These findings suggest that the ecosystem might be vulnerable to external perturbations targeting highly connected species, although other properties might provide resilience and resistance, resulting in a rearranged structure that preserves its original functions. Furthermore, we identified several species as important in terms of different aspects of trophic structure and functioning, and response to perturbations. We suggest that generalist species, mainly fishes, play a crucial role in the ecosystem's bentho-pelagic coupling process and that other species, besides the longtail southern cod *Patagonotothen ramsayi* and the Fueguian sprat *Sprattus fuegensis*, should be considered as relevant energy transfers for the ecosystem. Finally, we argue that the diversity of species, including both the benthic and pelagic habitats, is responsible for seccuring the connectivity within the food web against perturbations, therefore contributing to the structure and stability of the ecosystem.

@@

In terms of complexity, the MPA Namuncurá - Burdwood Bank food web consisted of 1788 predator-prey interactions and 380 species, where 93% of them were defined at the species taxonomical level (Figure 2, Table S2). The food web presented a link density (e.g., the average number of interactions per species) of 4.70, and a connectance of 0.01. Almost half of the consumers were omnivores (0.48), feeding on sources at different trophic levels. The food web displayed a small-world pattern, meaning that the path length was lower and the clustering coefficient higher than the random networks (Table 2).

@@

Considering the topological role, 'module specialist' species were the most frequent and presented a wide TL range (1 - 4.78), as well as 'module hub' species (TL = 1 - 3.9); 'module connector' was constrained to mid-TLs (2 - 3.86); and 'network connector', was represented by only one trophic species: detritus (Figure 4D, see Figure S2 for species' topological roles in a food web graph framework). Here it is important to highlight the two latter topological roles because they are responsible for linking modules and maintaining the connectivity of the food web: 42 species (1 network connector + 41 module connectors) from 19 different functional groups with a TL range = 1 - 3.86. The 41 species with a module connector role represented these functional groups: Amphipoda, Bivalvia, Brachiopoda, Bryozoa, Hydrozoa (as 'Cnidaria_benthic'), Copepoda, Cumacea, Decapoda, Echinodermata, fish (bentho-pelagic and demersal Osteichthyes, and Chondrychthyes), Foraminifera, Polychaeta, Porifera, Pycnogonida (as 'Benthos_Misc') and zooplankton (see Supplementary Material Table S3 for the identity of the species).

@@

```{r figure4, echo=FALSE, fig.width = 6.5, fig.height = 6.5, fig.align='center', fig.cap="Species-level properties by trophic level: (A) betweenness, (B) closeness, (C) trophic similarity, and (D) topological role. Each point represents a species. Linear regressions for betweenness ($y = 74.82x - 116.82, R^2 = 0.05, p-value < 0.01$), closeness ($y = 1.04e-05x - 9.25e-4, R^2 = 0.004, p-value = 0.11$) and trophic similarity ($y = -0.02x + 0.1 , R^2 = 0.07, p-value < 0.01$). Note that for panels A, B and C only species with TLs equal or greater than 2 were considered."}
# Betweenness, Closeness, Trophic similarity & Topological role by Trophic level

>>>>

abstract: |
  Globally, the total area of the world ocean designated under marine protection has increased in the recent decades. The majority of these Marine Protected Areas (MPAs) focus on the presence of particularly vulnerable, keystone, or charismatic species, large numbers of endemic species, and/or high biodiversity across taxonomic levels. In the sub-Antarctic region, the level of ocean protection is mainly associated to oceanic islands, except for the MPAs Namuncurá - Burdwood Bank I and II (MPA N-BB, ~53º–55ºS and ~56º–62ºW), which are associated to a submarine plateau and its southern adjacent deep slope region, respectively. Here, we present the first detailed analysis of the network of predator-prey interactions (food web) for the MPA N-BB ecosystem. We applied a network approach to characterise the food web in terms of complexity and structure, and identifying the species' role in such a framework. In terms of complexity, the MPA N-BB food web consisted of 1788 predator-prey interactions and 379 species, with a link density of 4.72 and a connectance of 0.01. In terms of structure, almost half of the consumers were omnivores (0.48), and the network displayed a small-world pattern. These findings suggest that the ecosystem might be vulnerable to external perturbations targeting highly connected species, although other properties might provide resilience and resistance, resulting in a rearranged structure that preserves its original functions. Furthermore, we identified several species as important in terms of different aspects of trophic structure and functioning, and response to perturbations. We suggest that generalist species, mainly fishes, play a crucial role in the ecosystem's bentho-pelagic coupling process and that other species, besides the longtail southern cod *Patagonotothen ramsayi* and the Fueguian sprat *Sprattus fuegensis*, should be considered as relevant energy transfers for the ecosystem. Finally, we argue that the diversity of species, including both the benthic and pelagic habitats, is responsible for seccuring the connectivity within the food web against perturbations, therefore contributing to the structure and stability of the ecosystem.

@@

In terms of complexity, the MPA Namuncurá - Burdwood Bank food web consisted of 1788 predator-prey interactions and 379 species, where 93% of them were defined at the species taxonomical level (Figure 2, Table S2). The food web presented a link density (e.g., the average number of interactions per species) of 4.72, and a connectance of 0.01. Almost half of the consumers were omnivores (0.48), feeding on sources at different trophic levels. The food web displayed a small-world pattern, meaning that the path length was lower and the clustering coefficient higher than the random networks (Table 2).

@@

Considering the topological role, 'module specialist' species were the most frequent and presented a wide TL range (1 - 4.78), as well as 'module hub' species (TL = 1 - 3.92); 'module connector' was constrained to mid-TLs (2 - 3.86); and 'network connector', was represented by only one trophic species: detritus (Figure 4D, see Figure S2 for species' topological roles in a food web graph framework). Here it is important to highlight the two latter topological roles because they are responsible for linking modules and maintaining the connectivity of the food web: 42 species (1 network connector + 41 module connectors) from 19 different functional groups with a TL range = 1 - 3.86. The 41 species with a module connector role represented these functional groups: Amphipoda, Bivalvia, Brachiopoda, Bryozoa, Hydrozoa (as 'Cnidaria_benthic'), Copepoda, Cumacea, Decapoda, Echinodermata, fish (bentho-pelagic and demersal Osteichthyes, and Chondrychthyes), Foraminifera, Polychaeta, Porifera, Pycnogonida (as 'Benthos_Misc') and zooplankton (see Supplementary Material Table S3 for the identity of the species).

@@

```{r figure4, echo=FALSE, fig.width = 6.5, fig.height = 6.5, fig.align='center', fig.cap="Species-level properties by trophic level: (A) betweenness, (B) closeness, (C) trophic similarity, and (D) topological role. Each point represents a species. Linear regressions for betweenness ($y = 74.97x - 117.35, R^2 = 0.05, p-value < 0.01$), closeness ($y = 9.33e-06x - 9.31e-4, R^2 = 0.003, p-value = 0.15$) and trophic similarity ($y = -0.02x + 0.11 , R^2 = 0.07, p-value < 0.01$). Note that for panels A, B and C only species with TLs equal or greater than 2 were considered."}
# Betweenness, Closeness, Trophic similarity & Topological role by Trophic level
